---
title: "ACLU Project"
author: "Chris Kaiser-Nyman"
date: "5/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r chunk_name, results="hide"}
library(tidyverse)
set.seed(101)
```




Load data:
```{r}

data <- read_csv("aclu_Nov_2014.csv")
```




Change column names so spaces are underscores:

```{r}

names(data) <- make.names(names(data),unique = TRUE)
names(data) <- gsub(x = names(data), pattern = "\\.\\.", replacement = "_")  
names(data) <- gsub(x = names(data), pattern = "\\.", replacement = "_")

```


Creating a column for total jail time for each case:

```{r}
data$total_jail_time <- sum(data$Jail_Time_1_1_,
                            data$Jail_Time_2_1_,
                            data$Jail_Time_3_1_,
                            data$Jail_Time_4_1_,
                            data$Jail_Time_5_1_,
                            data$Jail_Time_1_2_,
                            data$Jail_Time_2_2_,
                            data$Jail_Time_3_2_,
                            data$Jail_Time_4_2_,
                            data$Jail_Time_5_2_,
                            data$Jail_Time_1_3_,
                            data$Jail_Time_2_3_,
                            data$Jail_Time_3_3_,
                            data$Jail_Time_4_3_,
                            data$Jail_Time_5_3_,
                            data$Jail_Time_1_4_,
                            data$Jail_Time_2_4_,
                            data$Jail_Time_3_4_,
                            data$Jail_Time_4_4_,
                            data$Jail_Time_5_4_)
```







##########################################################
Merging text descriptions of charges
##########################################################



Charge code codebook was imported from here on April 5, 2019: https://oag.ca.gov/law/code-tables
Txt link: https://oag.ca.gov/sites/all/files/agweb/law-enforcement/code-tables/macrcode.txt?051520190840
Direct csv link: https://oag.ca.gov/sites/all/files/agweb/law-enforcement/code-tables/macrcode.csv?051520190840
```{r}

chargecodes <- read_csv("chargecodes.csv", col_names = FALSE)
# name columns
# it isn't clear to me what columns 2 and 3 are for
names(chargecodes) <- c("BCS_code", "no_2","no_3", "m_f", "offense_number", "penal_code", "offense", "sentence")

```



Clean charge codebook:
```{r}

# create key
chargecodes$key <- paste(chargecodes$penal_code, chargecodes$offense_number, chargecodes$m_f, sep="")
# remove any spaces so keys will line up
chargecodes$key <- gsub(" ", "", chargecodes$key, fixed = TRUE)
# make everything uppercase
chargecodes$key <- toupper(chargecodes$key)

# Some charge codes have multiple descriptions :( let's paste 
# those together, so each charge code has a unique description
charge_codebook <- chargecodes %>%
  group_by(key) %>%
  summarise(unique_description = paste(offense, collapse = "----,---- "))

#Manually enter some of the highest-frequency ones into the chargecodes data table key
line1 <- c("PC459/460(A)F", "First or second degree burglary / Felony BURGLARY:FIRST DEGREE")
line2 <- c("PC211/212.5(C)F", "Robbary 1st or 2nd degree / Felony ROBBERY:SECOND DEGREE")
line3 <- c("PC484/488M", "PC484 is a lot of things / PC488(M) is PETTY THEFT")
line4 <- c("PC459/460(B)F", "Shoplifting or 1st or 2nd degree Burglary/ Felony BURGLARY:SECOND DEGREE")
line5 <- c("PC422F", "PC 422 (F) could be VIOLATE CIVIL RIGHTS:CAUSE VIOLENT INJURY/ETC, or CAUSE PROPERTY DAMAGE, or THREATEN CRIME WITH INTENT TO TERRORIZE")
line6 <- c("PC496DF", "ATTEMPT TO COMMIT 496 PC:STOLEN PROPERTY")
line7 <- c("HS11377M", "HS 11377(A) is POSSESS CONTROLLED SUBSTANCE, but this has no (A)")
line8 <- c("PC3455F", "Felony Violation of Post-Release Community Supervision (PRCS)")
line9 <- c("PC3000.08F",  "Violation of `Sex-Offender` Supervision")

# update the codebook
charge_codebook <- rbind(charge_codebook, line1, line2, line3, line4, line5, line6, line7, line8, line9)

# remove the manual update objects from the workspace
rm(line1, line2, line3, line4, line5, line6, line7, line8, line9)

```


Create BCS Codebook
```{r}

bcs_codebook <- chargecodes %>%
  group_by(key) %>%
  summarise(BCS_code_added = paste(BCS_code, collapse = "----,---- "))
# issue here is that a few BCS codes are both misdemeanor (M) and felony (F), 
# so not sure how severtiy is measured for those

```



##############################################################
######## Merge Charge Descriptions with Main Dataframe #######
##############################################################


Create a function that cleans up the Charge Code Columns in the main dataframe

```{r}

charge_cleaner <- function(df, column, severity) {
  # these two lines allow the function to call the columns without quotes
  # column <- deparse(substitute(column))
  # severity <- deparse(substitute(severity))
  
  # remove ampersands (&)
  df[[column]] <- sub("&", "", df[[column]])
  # change CVC to VC
  df[[column]] <- sub("CVC", "VC", df[[column]])

  # create key
  df[[paste(column, "key", sep = "_")]] <- paste(df[[column]], df[[severity]], sep="")
  # remove spaces
  df[[paste(column, "key", sep = "_")]] <- gsub(" ", "", df[[paste(column, "key", sep = "_")]], fixed = TRUE)
  # make uppercase
  df[[paste(column, "key", sep = "_")]] <- toupper(df[[paste(column, "key", sep = "_")]])
  
  # turn "NANA" into actual NAs
  df[[paste(column, "key", sep = "_")]] <- na_if(df[[paste(column, "key", sep = "_")]], "NANA")
  
  # select new key column
  as.data.frame(df[,c(1, ncol(df))])
}



```


Apply this function over all the charge columns

```{r}

data_charge_list <- pmap(list(df = list(data),
              column = list("Charge_Code_1", "Charge_Code_2", "Charge_Code_3",
                            "Charge_Code_4",  "Charge_Code_5", "Charge_Code_6", 
                            "Charge_Code_7", "Charge_Code_8", "Charge_Code_9", "Charge_Code_10", 
                            "Charge_Code_11"),
              severity = list("Severity_1_", "Severity_2_", "Severity_3_", "Severity_4_", 
                              "Severity_5_", "Severity_6_", "Severity_7_", "Severity_8_", 
                              "Severity_9_", "Severity_10_", "Severity_11_")),
              charge_cleaner)

```



Create dataframe of these charge codes

```{r}
justkeys <- as.data.frame(cbind(data_charge_list[[1]][[2]], data_charge_list[[2]][[2]], 
                                data_charge_list[[3]][[2]], data_charge_list[[4]][[2]],
                                data_charge_list[[5]][[2]], data_charge_list[[6]][[2]], 
                                data_charge_list[[7]][[2]], data_charge_list[[8]][[2]],
                                data_charge_list[[9]][[2]], data_charge_list[[10]][[2]],
                                data_charge_list[[11]][[2]]))
```


Create "id" column for sorting later

```{r}
justkeys$id <- 1:nrow(justkeys)
```


Create function for merging the keys and the charge descriptions

```{r}
charge_merge <- function(column) {
  merge(x = justkeys, 
        y = charge_codebook, 
        by.x = column, 
        by.y = "key", 
        all.x = TRUE) %>%
  select(id, unique_description)
}
```



Map it over all the "justkeys" columns

```{r}
charge_dscrpt <- map(c("V1", "V2", "V3", "V4", "V5", "V6", 
                       "V7", "V8", "V9", "V10", "V11"), charge_merge)
```


Sort the output so they'll line up when merged with main dataframe

```{r}
charge_dscrpt[[1]] <- arrange(charge_dscrpt[[1]], id)
charge_dscrpt[[2]] <- arrange(charge_dscrpt[[2]], id)
charge_dscrpt[[3]] <- arrange(charge_dscrpt[[3]], id)
charge_dscrpt[[4]] <- arrange(charge_dscrpt[[4]], id)
charge_dscrpt[[5]] <- arrange(charge_dscrpt[[5]], id)
charge_dscrpt[[6]] <- arrange(charge_dscrpt[[6]], id)
charge_dscrpt[[7]] <- arrange(charge_dscrpt[[7]], id)
charge_dscrpt[[8]] <- arrange(charge_dscrpt[[8]], id)
charge_dscrpt[[9]] <- arrange(charge_dscrpt[[9]], id)
charge_dscrpt[[10]] <- arrange(charge_dscrpt[[10]], id)
charge_dscrpt[[11]] <- arrange(charge_dscrpt[[11]], id)
```


Merge them into the main dataframe

```{r}
data2 <- cbind(data, charge_dscrpt[[1]][[2]], charge_dscrpt[[2]][[2]], charge_dscrpt[[3]][[2]], 
               charge_dscrpt[[4]][[2]], charge_dscrpt[[5]][[2]], charge_dscrpt[[6]][[2]], 
               charge_dscrpt[[7]][[2]], charge_dscrpt[[8]][[2]], charge_dscrpt[[9]][[2]], 
               charge_dscrpt[[10]][[2]], charge_dscrpt[[11]][[2]])
```



Rename these new charge description columns:

```{r}
oldnames = c("charge_dscrpt[[1]][[2]]", "charge_dscrpt[[2]][[2]]", "charge_dscrpt[[3]][[2]]",
             "charge_dscrpt[[4]][[2]]", "charge_dscrpt[[5]][[2]]", "charge_dscrpt[[6]][[2]]",
             "charge_dscrpt[[7]][[2]]", "charge_dscrpt[[8]][[2]]", "charge_dscrpt[[9]][[2]]",
             "charge_dscrpt[[10]][[2]]", "charge_dscrpt[[11]][[2]]")
newnames = c("Charge_1", "Charge_2", "Charge_3", "Charge_4", "Charge_5", "Charge_6", 
             "Charge_7", "Charge_8", "Charge_9", "Charge_10", "Charge_11")

data2 <- data2 %>%
  rename_at(vars(oldnames), ~ newnames)
```






##############################################################
########      Merge BCS Codes with Main Dataframe      #######
##############################################################


Create function for matching BCS codes to the key column

```{r}
bcs_merge <- function(column) {
  merge(x = justkeys, 
        y = bcs_codebook, 
        by.x = column, 
        by.y = "key", 
        all.x = TRUE) %>%
  select(id, BCS_code_added)
}
```



Map it over all the "justkeys" columns

```{r}
bcs_list <- map(c("V1", "V2", "V3", "V4", "V5", "V6", 
                       "V7", "V8", "V9", "V10", "V11"), bcs_merge)
```


Sort the output so they will line up when merged

```{r}
bcs_list[[1]] <- arrange(bcs_list[[1]], id)
bcs_list[[2]] <- arrange(bcs_list[[2]], id)
bcs_list[[3]] <- arrange(bcs_list[[3]], id)
bcs_list[[4]] <- arrange(bcs_list[[4]], id)
bcs_list[[5]] <- arrange(bcs_list[[5]], id)
bcs_list[[6]] <- arrange(bcs_list[[6]], id)
bcs_list[[7]] <- arrange(bcs_list[[7]], id)
bcs_list[[8]] <- arrange(bcs_list[[8]], id)
bcs_list[[9]] <- arrange(bcs_list[[9]], id)
bcs_list[[10]] <- arrange(bcs_list[[10]], id)
bcs_list[[11]] <- arrange(bcs_list[[11]], id)
```


Merge them into the main dataframe

```{r}
data3 <- cbind(data2, bcs_list[[1]][[2]], bcs_list[[2]][[2]], bcs_list[[3]][[2]], 
               bcs_list[[4]][[2]], bcs_list[[5]][[2]], bcs_list[[6]][[2]], 
               bcs_list[[7]][[2]], bcs_list[[8]][[2]], bcs_list[[9]][[2]], 
               bcs_list[[10]][[2]], bcs_list[[11]][[2]])
```



Rename these new BCS code columns:

```{r}
oldnames = c("bcs_list[[1]][[2]]", "bcs_list[[2]][[2]]", "bcs_list[[3]][[2]]",
             "bcs_list[[4]][[2]]", "bcs_list[[5]][[2]]", "bcs_list[[6]][[2]]",
             "bcs_list[[7]][[2]]", "bcs_list[[8]][[2]]", "bcs_list[[9]][[2]]",
             "bcs_list[[10]][[2]]", "bcs_list[[11]][[2]]")
newnames = c("bcs_charge_1", "bcs_charge_2", "bcs_charge_3", "bcs_charge_4", 
             "bcs_charge_5", "bcs__charge_6", "bcs_charge_7", "bcs_charge_8", 
             "bcs_charge_9", "bcs_charge_10", "bcs_charge_11")

data3 <- data3 %>%
  rename_at(vars(oldnames), ~ newnames)
```





##############################################################
########      Create clean demographic columns      #########
##############################################################



Create new columns for race, gender, zip codes, and cities, including only the most common instances of each:

```{r}
data3 <- data3 %>%
  mutate(clean_race = recode(Race, 'F' = 'Other',
                           'O' = 'Other' ,
                           'X' = 'Other' ,
                           'L' = 'Other' ,
                           'C' = 'Other' ,
                           'I' = 'Other' ,
                           'P' = 'Other' ,
                           'Z' = 'Other' ,
                           'S' = 'Other' ,
                           'V' = 'Other' ,
                           'G' = 'Other' ,
                           'J' = 'Other' ,
                           'U' = 'Other'),
         clean_gender = recode(Gender, 'P' = NA_character_, "U" = NA_character_),
         common_zips = recode(Zip_Code, `0`	=	0, 
                              `94509`	=	94509,
                              `94565`	=	94565,
                              `94806`	=	94806,
                              `94804`	=	94804,
                              `94520`	=	94520,
                              `94801`	=	94801,
                              `94553`	=	94553,
                              `94561`	=	94561,
                              `45650`	=	45650,
                              `94513`	=	94513,
                              `45090`	=	45090,
                              `48060`	=	48060,
                              `94521`	=	94521,
                              `94519`	=	94519,
                              `94803`	=	94803,
                              `94523`	=	94523,
                              .default = 99999),
         common_cities = recode(City, "Antioch"	=	"Antioch",
                              "Richmond"	=	"Richmond",
                              "Pittsburg"	=	"Pittsburg",
                              "Concord"	=	"Concord",
                              "San Pablo"	=	"San Pablo",
                              "Martinez"	=	"Martinez",
                              "Bay Point"	=	"Bay Point",
                              "Oakland"	=	"Oakland",
                              "Oakley"	=	"Oakley",
                              "Brentwood"	=	"Brentwood",
                              "Vallejo"	=	"Vallejo",
                              "San Francisco"	=	"San Francisco",
                              "Walnut Creek"	=	"Walnut Creek",
                              "El Sobrante"	=	"El Sobrante",
                              "Pleasant Hill"	=	"Pleasant Hill",
                              "Pinole"	=	"Pinole",
                              "Hercules"	=	"Hercules",
                              "Rodeo"	=	"Rodeo",
                              .default = "other"))
```




Write this out to a csv:

```{r}

write_csv(data3, "ccc_final_data.csv")

```














I found frequency tables for the columns very helpful in determining what my data looked like at any given time. Below I have copied the very repetitve code I wrote to do this, but it should be able to be done as a function without too much trouble. Having it map over ALL the columns in the dataframe could also be helpful.


```{r}

bcs1table <- as.data.frame(table(data3$bcs_charge_1, useNA = "ifany")) %>%
  arrange(-Freq)

racetable <- as.data.frame(table(data3$Race, useNA = "ifany")) %>%
  arrange(-Freq)

gendertable <- as.data.frame(table(data3$Gender, useNA = "ifany")) %>%
  arrange(-Freq)

citytable <- as.data.frame(table(data3$City, useNA = "ifany")) %>%
  arrange(-Freq)

counttable <- as.data.frame(table(data3$Count_1_, useNA = "ifany")) %>%
  arrange(-Freq)

chargetable  <- as.data.frame(table(data3$Charge_Code_1, useNA = "ifany")) %>%
  arrange(-Freq)

severitytable <- as.data.frame(table(data3$Severity_1_, useNA = "ifany")) %>%
  arrange(-Freq)

dispositiontable  <- as.data.frame(table(data3$Disposition_1_, useNA = "ifany")) %>%
  arrange(-Freq)

prisontable <- as.data.frame(table(data3$Prison_1_1_, useNA = "ifany")) %>%
  arrange(-Freq)

enhancetable <- as.data.frame(table(data3$Charge_Enhance_1_1_, useNA = "ifany")) %>%
  arrange(-Freq)

probationtable <- as.data.frame(table(data3$Probation_Rev_Type_1_1_, useNA = "ifany")) %>%
  arrange(-Freq)

defenhancetable <- as.data.frame(table(data3$Defendant_Enhance_1, useNA = "ifany")) %>%
  arrange(-Freq)

incidentagencytable <- as.data.frame(table(data3$Incident_Agency_1, useNA = "ifany")) %>%
  arrange(-Freq)

amounttable <- as.data.frame(table(data3$Amount, useNA = "ifany")) %>%
  arrange(-Freq)

acttable <- as.data.frame(table(data3$Act_1, useNA = "ifany")) %>%
  arrange(-Freq)

restitutiontable <- as.data.frame(table(data3$Restitution, useNA = "ifany")) %>%
  arrange(-Freq)
```



We can put these into an excel file:
```{r}
# we can then put these into a excel file with each dataframe as a separate sheet
# requires 
library(openxlsx)

# Create a blank workbook
OUT <- createWorkbook()

# Add some sheets to the workbook
addWorksheet(OUT, "bcs1")
addWorksheet(OUT, "race")
addWorksheet(OUT, "city")
addWorksheet(OUT, "count")
addWorksheet(OUT, "charge")
addWorksheet(OUT, "severity")
addWorksheet(OUT, "disposition")
addWorksheet(OUT, "prison")
addWorksheet(OUT, "charge enhance")
addWorksheet(OUT, "probation")
addWorksheet(OUT, "defendant enhance")
addWorksheet(OUT, "incident agency")
addWorksheet(OUT, "amount")
addWorksheet(OUT, "act")
addWorksheet(OUT, "restitution")

# Write the data to the sheets
writeData(OUT, sheet = "bcs1", x = bcs1table)
writeData(OUT, sheet = "race", x = racetable)
writeData(OUT, sheet = "city", x = citytable)
writeData(OUT, sheet = "count", x = counttable)
writeData(OUT, sheet = "charge", x = chargetable)
writeData(OUT, sheet = "severity", x = severitytable)
writeData(OUT, sheet = "disposition", x = dispositiontable)
writeData(OUT, sheet = "prison", x = prisontable)
writeData(OUT, sheet = "charge enhance", x = enhancetable)
writeData(OUT, sheet = "probation", x = probationtable)
writeData(OUT, sheet = "defendant enhance", x = defenhancetable)
writeData(OUT, sheet = "incident agency", x = incidentagencytable)
writeData(OUT, sheet = "amount", x = amounttable)
writeData(OUT, sheet = "act", x = acttable)
writeData(OUT, sheet = "restitution", x = restitutiontable)

# Export the file
saveWorkbook(OUT, "summary_tables.xlsx")


```



Obviously summary tables can be created and added to the excel file as tabs for any columns you wish















